Phylogenetic clustering with a Markov-modulated Poisson process

### 2015-10-18 doc/stats-seminar/presentation.tex
Almost finished stats seminar talk...

\begin{frame}{Making an HIV phylogeny}
    \begin{tikzpicture}

    \end{tikzpicture}
\end{frame}

@@

\section{Challenges: validation}

@@

\begin{frame}
    \frametitle{Genetic distance based clustering}
    \tikz[remember picture,overlay] \node [anchor=north east, inner sep=1.5 cm] at (current page.north east) {\includegraphics[width=0.3\textwidth]{timetree-riskgroup-legend}};

    \tikz[remember picture,overlay] \node [anchor=north west, inner ysep=1 cm] at (current page.north west) {\includegraphics[width=0.6\textwidth]{timetree-pd-polar}};

    \tikz[remember picture,overlay] \node [anchor=south east, inner ysep=0 cm] at (current page.south east) {\includegraphics[width=0.6\textwidth]{timetree-riskgroup-polar}};
\end{frame}

\begin{frame}
    \frametitle{Model-based clustering}
    \tikz[remember picture,overlay] \node [anchor=north east, inner sep=1.5 cm] at (current page.north east) {\includegraphics[width=0.3\textwidth]{timetree-riskgroup-legend}};

    \tikz[remember picture,overlay] \node [anchor=north west, inner ysep=1 cm] at (current page.north west) {\includegraphics[width=0.6\textwidth]{timetree_4}};

    \tikz[remember picture,overlay] \node [anchor=south east, inner ysep=0 cm] at (current page.south east) {\includegraphics[width=0.6\textwidth]{timetree-riskgroup-polar}};
\end{frame}

\begin{frame}
    \frametitle{Preliminary comparison}
    \tikz[remember picture,overlay] \node [anchor=north east, inner ysep=1.5 cm, inner xsep=0.5 cm, text width=6cm] at (current page.north east) {$\longleftarrow$ 79 clusters \\ median [IQR] 29 [16-56] individuals};

    \tikz[remember picture,overlay] \node [anchor=north west, inner ysep=1 cm] at (current page.north west) {\includegraphics[width=0.6\textwidth]{cluster4}};

    \tikz[remember picture,overlay] \node [anchor=south west, inner xsep=0.5 cm, inner ysep=1 cm, text width=6cm] at (current page.south west) {\hfill 221 clusters $\longrightarrow$ \\ \hfill median [IQR] 5 [8-15] individuals};

    \tikz[remember picture,overlay] \node [anchor=south east, inner ysep=0 cm] at (current page.south east) {\includegraphics[width=0.6\textwidth]{timetree-pd-polar}};
\end{frame}

\begin{frame}
    \frametitle{Challenges}
    \tikz[remember picture,overlay] \node [anchor=north east, inner ysep=1.5 cm, inner xsep=0.5 cm, text width=6.5cm] at (current page.north east) {$\longleftarrow$ likelihood ratio test doesn't work? \\ $\qquad$ how to define clusters?};

    \tikz[remember picture,overlay] \node [anchor=north west, inner ysep=1 cm] at (current page.north west) {\includegraphics[width=0.6\textwidth]{timetree_6}};

    \tikz[remember picture,overlay] \node [anchor=south west, inner xsep=1.5 cm, inner ysep=1 cm] at (current page.south west) {scaling tip branches $\longrightarrow$};

    \tikz[remember picture,overlay] \node [anchor=south east, inner ysep=0 cm] at (current page.south east) {\includegraphics[width=0.6\textwidth]{timetree-qnorm_3}};
\end{frame}

\begin{frame}
    \frametitle{Acknowledgements}

    \textbf{BC Centre for Excellence in HIV/AIDS} \\
    $\qquad$ Art Poon \\
    $\qquad$ Richard Liang \\
    $\qquad$ Jeffrey Joy \\
    $\qquad$ T. Nguyen \\

    \vspace{3cm}
    \includegraphics[width=0.2\textwidth]{logos/nserc}
    \includegraphics[width=0.35\textwidth]{logos/cfe}
    \includegraphics[width=0.35\textwidth]{logos/btp}
    \includegraphics[width=0.1\textwidth]{logos/ubc}
\end{frame}

>>>>

\begin{frame}{Population-level HIV sequencing}
    \begin{tikzpicture}
        \node (a) {\includegraphics[height=1.5cm]{person7}};
        \node (b) [right=of a] {\includegraphics[height=1.5cm]{blood_testing}};
        \draw [->, >=stealth, thick] (a) -- (b);
        \uncover<2->{
        \node (c) [below=of b] {\includegraphics[height=1.5cm]{hiv}\includegraphics[height=1.5cm]{alignment}};
        \draw [->, >=stealth, thick] (b) -- (c);
        }
        \uncover<3->{
        \node (d) [right=of c, rectangle, draw, inner ysep=12pt] {\includegraphics[height=1.5cm]{medical}};
        \draw [->, >=stealth, thick] (c) -- (d);
        }
        \uncover<4->{
        \node (e) [below=of c] {\includegraphics[height=0.25cm]{sequence1}};
        \draw [->, >=stealth, thick] (c) -- (e);
        }
    \end{tikzpicture}
\end{frame}

\begin{frame}{Maximum likelihood HIV phylogeny}
    \begin{columns}
        \begin{column}{0.5\textwidth}
            \includegraphics[height=1cm]{person7}\includegraphics[height=0.25cm]{sequence1}

            \includegraphics[height=1cm]{person8}\includegraphics[height=0.25cm]{sequence2}

            \includegraphics[height=1cm]{person9}\includegraphics[height=0.25cm]{sequence4}

            \includegraphics[height=1cm]{person10}\includegraphics[height=0.25cm]{sequence5}
        \end{column}
        \begin{column}{0.5\textwidth}
            \centering
            \uncover<2->{
            \includegraphics[width=\textwidth]{unrooted}

            FastTree 2
            }
        \end{column}
    \end{columns}

    \tiny
    Price, Morgan N., Paramvir S. Dehal, and Adam P. Arkin. ``FastTree
    2 -- approximately maximum-likelihood trees for large alignments.'' PloS one
    5.3 (2010): e9490.\par
\end{frame}

\begin{frame}{Time scaling the phylogeny}

    \begin{columns}
        \begin{column}{0.5\textwidth}
            \centering
            \includegraphics[height=1cm]{unrooted}

            \includegraphics[width=\textwidth]{rtt}
        \end{column}
        \begin{column}{0.5\textwidth}
            \uncover<2->{
            \includegraphics[width=\textwidth]{rooted}
            }
        \end{column}
    \end{columns}

    \tiny
    Drummond, Alexei, G. Oliver, and Andrew Rambaut. ``Inference of viral
    evolutionary rates from molecular sequences.'' Advances in parasitology 54
    (2003): 331-358.\par
\end{frame}

@@

\section{Challenges}

\section{Previous projects}

\begin{frame}{Phylogenetic estimation of transmitted/founder viruses}
    \begin{center}
    \begin{tikzpicture}
        [scale=1.7, every node/.style={inner sep=0pt}]
        \node (root) at (1, 2) {\includegraphics[width=1cm]{hiv}};
        \node (lc) at (0.5, 1) {\includegraphics[width=1cm]{hiv}};
        \node (t1) at (0, 0) {\includegraphics[width=1cm]{hiv}};
        \node (t2) at (1, 0) {\includegraphics[width=1cm]{hiv}};
        \node (t3) at (2, 0) {\includegraphics[width=1cm]{hiv}};
        \draw (root) -- (lc) -- (t1);
        \draw (lc) -- (t2);
        \draw (root) -- (t3);

        \node at (3, 2) {$\longrightarrow$ transmitted/founder};
        \node at (3, 0) {$\longrightarrow$ sampled};
    \end{tikzpicture}
\end{center}

    Find patients with known early sequences; try to reconstruct with
    phylogenetic methods.

    \tiny
    McCloskey, Rosemary M., et al. ``An Evaluation of Phylogenetic Methods for
    Reconstructing Transmitted HIV Variants using Longitudinal Clonal HIV
    Sequence Data.'' Journal of virology 88.11 (2014): 6181-6194.\par
\end{frame}

\begin{frame}{Phylogenetic estimation of transmitted/founder viruses}
    \begin{description}
        \setlength{\itemsep}{16pt}
        \item [{\Huge55}] publications with early and late sequences
        \item [{\Huge232}] patients
        \item [{\Huge335}] gene data sets
        \item [{\Huge19,486}] total HIV sequences
    \end{description}
\end{frame}

\begin{frame}{Phylogenetic estimation of transmitted/founder viruses}
    \includegraphics[width=\textwidth]{mccloskey2014evaluation_f1}
\end{frame}

\begin{frame}{Phylogenetic methods are informative}
    \includegraphics[height=\textheight]{mccloskey2014evaluation_f2}
\end{frame}

\begin{frame}{Single origin of Hepatitis C virus Q80K polymorphism}
    \begin{itemize}
        \setlength{\itemsep}{12pt}
        \item The Q80K polymorphism occurs in 5-47\% of HCV cases
        \item Confers mild resistance to some therapies 
        \item We investigated the origin and transmissibility of Q80K using
            publicly available sequence data
    \end{itemize}

    \vspace{1cm}
    \tiny Lin, M. Valerie, and Raymond Chung. ``Recent FDA approval of sofosbuvir and simeprevir. Implications for current HCV treatment.'' Clinical Liver Disease 3.3 (2014): 65-68. \par
    \tiny McCloskey, Rosemary M., et al. ``Global origin and transmission of hepatitis C virus NS3 Q80 K polymorphism.'' Journal of Infectious Diseases (2014): jiu613. \par
\end{frame}

\begin{frame}{Single origin of Hepatitis C virus Q80K polymorphism}
    \begin{description}
        \setlength{\itemsep}{16pt}
        \item [{\Huge160 071}] unfiltered sequences
        \item [{\Huge13}] countries of origin
        \item [{\Huge4}] filtering steps
        \item [{\Huge724}] filtered sequences
    \end{description}
\end{frame}

\begin{frame}{Single origin of Hepatitis C virus Q80K polymorphism}
    \includegraphics[width=\textwidth]{mccloskey2015global_f1}

    \includegraphics[width=\textwidth]{mccloskey2015global_supp}
\end{frame}

\begin{frame}
    \frametitle{Thanks}
    \begin{columns}
        \begin{column}{0.5\textwidth}
            \textbf{BC Centre for Excellence in HIV/AIDS} \\
            $\qquad$ Art Poon \\
            $\qquad$ Richard Liang \\
            $\qquad$ P. Richard Harrigan \\
            $\qquad$ Jeffrey Joy \\
            $\qquad$ Thuy Nguyen
        \end{column}
        \begin{column}{0.5\textwidth}
            \textbf{UBC} \\
            $\qquad$ Sarah Otto \\
            $\qquad$ Alexandre Bouchard-C\^ot\'e \\
            $\qquad$ Wayne Maddison \\
            $\qquad$ Sara Mostafavi \\
            \textbf{SFU} \\
            $\qquad$ Ryan Morin \\
            $\qquad$ Marni Mishna \\
            $\qquad$ Cedric Chauve
        \end{column}
    \end{columns}

    \vspace{1cm}
    \includegraphics[width=0.35\textwidth]{cfe}
    \includegraphics[width=0.35\textwidth]{btp}
    \includegraphics[width=0.1\textwidth]{ubc}
    \hspace{0.5cm}
    \includegraphics[width=0.1\textwidth]{ubcstats}
\end{frame}

@@
